Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-08-27 Documents/Paper/ms.tex
More editing

If $N' > N$, i.e. the proposal is to add objects to the model,
the extra parameters needed, $\{\xx_i\}_{i=N+1}^{N'}$,
are drawn from their prior conditional on the current value of the
hyperparameters, i.e. the interim prior $p(\xx | \hyperparams)$.

If $N' < N$, i.e. the proposal is to remove objects from the model,
then $N - N'$ objects must be selected for removal. All
of the objects have the same probability of being selected for removal.

@@

A proposal
for an object involves transforming its parameters to $[0, 1]$ using $F$,
making the proposal in terms of $u$, and then transforming back.
Specifically, the proposal chooses
a new value $\xx_i'$ from the current value $\xx_i$ as follows:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
u_i' &:=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
\xx_i' &:=& G(u_i'; \hyperparams).
\end{eqnarray}
where $\delta_u$ is drawn from a heavy-tailed distribution.

@@

\subsection{Proposals that change the hyperparameters,
keeping the objects fixed}\label{sec:proposal3}
Another kind of proposal that we will consider is a proposal that keeps all of
the objects fixed in place (i.e. leaves $\{\xx_i\}$) unchanged, but changes
the hyperparameter(s) from their current value $\hyperparams$
to a new value $\hyperparams'$. The proposal for the hyperparameter(s) is chosen
to satisfy detailed balance with respect to $p(\hyperparams)$.

The overall Metropolis acceptance probability
for this kind of move, if sampling the prior (or the constrained prior of DNS)
must include the following factor:
\begin{eqnarray}
\frac{\prod_{i=1}^N p(\xx_i | \hyperparams')}{\prod_{i=1}^N p(\xx_i | \hyperparams)}
\label{eqn:acceptance_prob}
\end{eqnarray}
Since this proposal leaves the objects $\{\xx_i\}$ fixed, it will usually not
affect the value of the likelihood, and therefore the likelihood will not need
to be recomputed.

@@

Therefore, we allow an additional move that changes $\hyperparams$ to a new
value $\hyperparams'$, but rather than leaving the objects $\{\xx_i\}$ fixed, we
drag them so they represent the distribution $p(\xx | \hyperparams')$ rather than
$p(\xx|\hyperparams)$. We do this by making use of the transformation functions
$F(x; \hyperparams)$ and $G(x; \hyperparams)$ defined in Section~\ref{sec:proposal2}.

The ``dragging'' process works as follows, and must be carried out on
each object:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
\xx_i' &:=& G(\xx_i; \hyperparams')
\end{eqnarray}

>>>>

If $N' > N$, i.e. the proposal is to add objects to the model,
the extra parameters needed, $\{\xx_i\}_{i=N+1}^{N'}$,
are drawn from their prior conditional on the current value of the
hyperparameters, i.e. the interim prior $p(\xx | \hyperparams)$.
If $N' < N$, i.e. the proposal is to remove objects from the model,
then $N - N'$ objects must be selected for removal. All
of the objects have the same probability of being selected for removal.

@@

A proposal
for an object involves transforming its parameters to $[0, 1]$ using $F$,
making the proposal in terms of $u$, and then transforming back.
Specifically, the proposal chooses
a new value $\xx_i'$ from the current value $\xx_i$ as follows:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
u_i' &:=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
\xx_i' &:=& G(u_i'; \hyperparams).
\end{eqnarray}
where $\delta_u$ is drawn from a heavy-tailed distribution as in
Equation~\ref{eqn:heavytailed}.

@@

\subsection{Proposals that change the hyperparameters,
keeping the objects fixed}\label{sec:proposal3}
Another kind of proposal that we will consider is a proposal that keeps all of
the objects fixed in place (i.e. leaves $\{\xx_i\}$) unchanged, but changes
the hyperparameter(s) from their current value $\hyperparams$
to a new value $\hyperparams'$. The proposal for the hyperparameter(s) is chosen
to satisfy detailed balance with respect to $p(\hyperparams)$ and should be
heavy-tailed.

The overall Metropolis acceptance probability
for this kind of move, if sampling the prior (or the constrained prior of DNS)
must include the following factor:
\begin{eqnarray}
\frac{\prod_{i=1}^N p(\xx_i | \hyperparams')}{\prod_{i=1}^N p(\xx_i | \hyperparams)}
\label{eqn:acceptance_prob}
\end{eqnarray}
Since this proposal leaves the objects $\{\xx_i\}$ fixed, it will usually not
affect the value of the likelihood, and therefore the likelihood will not need
to be recomputed. Therefore this proposal is most useful for mixing the values
of the hyperparameters $\hyperparams$ when the target distribution is highly
constrained.

@@

Therefore, we allow an additional move that changes $\hyperparams$ to a new
value $\hyperparams'$, but rather than leaving the objects $\{\xx_i\}$ fixed,
they are ``dragged''
so they represent the distribution $p(\xx | \hyperparams')$ rather than
$p(\xx|\hyperparams)$. We do this by making use of the transformation functions
$F(x; \hyperparams)$ and $G(x; \hyperparams)$ defined in
Section~\ref{sec:proposal2}.

The ``dragging'' process works as follows, and must be carried out on
each object:
\begin{eqnarray}
u_i &:=& F(\xx_i; \hyperparams)\\
\xx_i' &:=& G(\xx_i; \hyperparams')
\end{eqnarray}
In other words, the objects' parameters are transformed to $[0,1]$ using the
current value of the hyperparameters, and then transformed back using the
proposed value of the hyperparameters so they represent the new interim prior
rather than the old one.
